\chapter{Literature Survey}
\lhead{\bfseries LITERATURE SURVEY}
\label{literature_review}


The literature survey introduces advanced perception systems, machine learning techniques, and cooperative control in the context of autonomous systems. It explores methodologies for lane detection, essential for intelligent systems in vehicles, highlighting the evolution from feature-based and model-based methods to sophisticated deep learning techniques. The review identifies the computational challenges of these approaches and proposes an innovative Iterative Tree Search algorithm for efficiency. It also delves into path following challenges in UAVs and the trade-offs in path detection and following techniques, focusing on computational constraints and the need for real-time processing. 

The discourse on machine learning emphasizes the role of \gls{rl}, especially Deep \gls{rl} in unpredictable driving scenarios, stressing the balance between learning efficacy and safety. 

The review touches on cybersecurity in autonomous vehicles, underscoring the research gap in control systems and the efforts made to mitigate cyber-attacks on vehicular systems, incorporating control loop resilience strategies.

A significant body of work has explored \gls{vc} in railway systems, emphasizing advanced control techniques like \gls{mpc} and \gls{ml}-based control. Research highlights the potential of combining machine learning with control algorithms to enhance performance and safety. Developing resilient and efficient \gls{vc} control algorithms remains a key area for future exploration.


\newpage





\section{ Cooperative Control}

\red{TODO}



